### 2024-09-17 base-rates.qmd
20240917 - update nomogram plots based on FPR

For instance, if the starting probability is 0.5% and the [likelihood ratio](#sec-diagnosticLikelihoodRatio) is 10 (e.g., sensitivity = .90, specificity = .91: $\text{likelihood ratio} = \frac{\text{sensitivity}}{1 - \text{specificity}} = \frac{.9}{1-.91} = 10$) from a positive test (i.e., [positive likelihood ratio](#sec-positiveLikelihoodRatio)), the updated probability is less than 5%, as depicted in @fig-probabilityNomogramLine.

@@

The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the sensitivity (SN) and specificity (SP) of the test.

@@

Here is a nomogram plot from [Example 1](#sec-nateSilverExample1) from @Silver2012 (where [specificity](#sec-specificity) equals 1 minus the [false positive rate](#sec-falsePositiveRate)):

```{r}
petersenlab::nomogrammer(
  pretestProb = .04,
  SN = .50,
  SP = 1 - .05
)
```

@@

```{r}
petersenlab::nomogrammer(
  pretestProb = .014,
  SN = .75,
  SP = 1 - .10
)
```

@@

```{r}
petersenlab::nomogrammer(
  pretestProb = .00005,
  SN = 1,
  SP = 1 - .00008
)
```

@@

```{r}
petersenlab::nomogrammer(
  pretestProb = .3846272,
  SN = 1,
  SP = 1 - .00008
)
```

>>>>

For instance, if the starting probability is 0.5% and the [likelihood ratio](#sec-diagnosticLikelihoodRatio) is 10 (e.g., [sensitivity](#sec-sensitivity) = .90, [specificity](#sec-specificity) = .91: $\text{likelihood ratio} = \frac{\text{sensitivity}}{1 - \text{specificity}} = \frac{.9}{1-.91} = 10$) from a positive test (i.e., [positive likelihood ratio](#sec-positiveLikelihoodRatio)), the updated probability is less than 5%, as depicted in @fig-probabilityNomogramLine.

@@

The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test.

@@

The function can also create a nomogram plot using the [sensitivity](#sec-sensitivity) (SN) and [false positive rate](#sec-falsePositiveRate) of the test.
Here is a nomogram plot from [Example 1](#sec-nateSilverExample1) from @Silver2012 (where [specificity](#sec-specificity) equals 1 minus the [false positive rate](#sec-falsePositiveRate)):

```{r}
petersenlab::nomogrammer(
  pretestProb = .04,
  SN = .50,
  FPR = .05
)
```

@@

```{r}
petersenlab::nomogrammer(
  pretestProb = .014,
  SN = .75,
  FPR = .10
)
```

@@

```{r}
petersenlab::nomogrammer(
  pretestProb = .00005,
  SN = 1,
  FPR = .00008
)
```

@@

```{r}
petersenlab::nomogrammer(
  pretestProb = .3846272,
  SN = 1,
  FPR = .00008
)
```
